Trajectory Landing Runway Assignment

### 2023-03-01 x_trj-apt-landing-association.qmd
more landing runway associated -- spotted wrong assignment due to ground coverage / not filtered out trajectory points ==> trajectory phase model!

```{r}
ggplot() +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) +
  geom_sf(data = baw236_landing_ls)
```

@@

```{r}
# spatial join (aka "overlay" with "within")
# to identify most likely ~ closest runway (extended centerline-string)
estimate_rwy <- function(.trj_pts, .rwy_ctrl_ls, .ctrl_offset = 1000){
  # stop if not
  tmp <- st_join( .rwy_ctrl_ls, .trj_pts
                 , join = st_is_within_distance, dist = .ctrl_offset) %>%
  st_drop_geometry() %>%        # keep just the "spatially joined" dataframe
  na.omit() %>%                 # eliminate no hits
  group_by(REF) %>%
  summarise(N = n(), .groups = "drop") %>%
  mutate( TOT   = sum(N, na.rm = TRUE)       # na.rm should not be necessary
         ,TRUST = N / TOT)                   # share in case several runways

return(tmp)
}
```

```{r}
baw236_final_approach <- baw236 |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 10, lon = LON, lat = LAT)
# baw236_landing_ls <- baw236_landing |> cast_latlon_to_ls(.drop_coord = FALSE, .group_var = ICAO24)
ggplot() +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) +
  geom_sf(data = baw236_final_approach |> cast_latlon_to_ls())
```

```{r}
estimate_rwy(rwy_centerline_ls, baw236_final_approach |> cast_latlon_to_pts())
```

Try to understand why / how the closeness is calculated.
Current implementation shows N=1 ?!? while ok, it should be multiple hits.

>>>>

```{r}
apt_arrival_plot <- ggplot() +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF))

apt_arrival_plot +
  geom_sf(data = baw236_landing_ls)
```

@@

```{r}
#' (Gu)estimate the arrival runway based on alignment of landing trajectory points and extended runway centerline
#' 
#'spatial join (aka "overlay" with "within")
# to identify most likely ~ closest runway (extended centerline-string)
#'
#' @param .trj_pts 
#' @param .rwy_ctrl_ls 
#' @param .ctrl_offset 
#'
#' @return
#' @export
#'
#' @examples
estimate_rwy <- function(.trj_pts, .rwy_ctrl_ls, .ctrl_offset = 1000){
  # stop if not
  tmp <- st_join( .rwy_ctrl_ls, .trj_pts
                 , join = st_is_within_distance, dist = .ctrl_offset) %>%
  st_drop_geometry() %>%        # keep just the "spatially joined" dataframe
 # na.omit() %>%                 # eliminate no hits
  group_by(REF) %>%
  summarise(N = n(), .groups = "drop") %>%
  mutate( TOT   = sum(N, na.rm = TRUE)       # na.rm should not be necessary
         ,TRUST = N / TOT)                   # share in case several runways

return(tmp)
}

#' Assign most likely arrival landing
#'
#' @param .rwy_hit_score 
#'
#' @return
#' @export
#'
#' @examples
#' #' \dontrun{
#' }
arrival_runway_assignment <- function(.rwy_hit_score){
  .rwy_hit_score |> 
    dplyr::arrange(desc(N)) |> 
    dplyr::mutate(
      COMMENT = paste0("landing runway: ", REF," trust-Level (hits: ", N, "): ", TRUST)) |> 
    dplyr::filter(row_number() == 1)
}
```

```{r}
final_approach_trj <- baw236 |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 10, lon = LON, lat = LAT)

final_approach_trj_pts <- final_approach_trj |> cast_latlon_to_pts()

landing_runway <- estimate_rwy(final_approach_trj_pts, rwy_centerline_ls) |> 
  arrival_runway_assignment()
landing_runway <- landing_runway |> mutate(LAT = 51.5, LON = -0.45)

apt_arrival_plot +
  geom_sf(data = final_approach_trj_pts) +
  labs(subtitle = landing_runway$COMMENT)
```
production pipeline

TO----------------DOOOOOOOOOOOOOOOOOOOOOOOOOOOOO

* multiple "landing trajectories"
* iterate over landing trajectories

```{r}
final_trjs <- baws |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 10, lon = LON, lat = LAT)

final_trjs_pts <- final_trjs |> cast_latlon_to_pts()

ldg_rwy <- final_trjs_pts |> group_by(ICAO24) |> 
  group_modify(.f = ~ estimate_rwy(.x, rwy_centerline_ls) |> arrival_runway_assignment())
ldg_rwy
```

WRONG RUNWAY ASSIGNMENT as filtering does not assure LAT/LONs on final!

```{r}
picks <-  c("400615")

apt_arrival_plot +
  geom_sf(data = final_trjs_pts |> filter(ICAO24 %in% picks))
```

```{r}
baw74_landing <- baws |> filter(ICAO24 == "400615") |>  
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 20, lon = LON, lat = LAT)
baw74_landing_ls <- baw74_landing |> cast_latlon_to_ls(.drop_coord = FALSE, .group_var = ICAO24)

apt_arrival_plot +
  geom_sf(data = baw74_landing_ls)
```

NEED FOR TRAJECTORY SEGMENTATION / PHASE ASSOCIATION

* trajectory = (time-)ordered set of 4D positions
* check
  + runway-end crossing  <-- arrival or departure
  + airport box          <-- arrival or departure
* other milestone
  + gate/6NM final
  + extended runway box (e.g. 15NM)
  + other characteristic milestones
* phase points
  + top of decent (/ climb)
  + ...
